Notes on modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-06-18 docs/mvfits_paper/mvf.tex
updating doc

\par The combination of these two ideas alone results in an algorithm prone to a deficit of stepwise methods: there is the very likely possibility that the algorithm would confuse correlated basis vectors (e.g. $x^2$ may be confused for $x^4$).

@@

  f(\vec{x}) = \muf + \sif \; \frac{ \sum_{r=0}^{R} a_r \, \phi_{r}(\vec{x}) }{ 1 - \sum_{v=0}^{V} b_v \, \phi_{v}(\vec{x}) } \;,

@@

where $\muf$ is the additive mean of $f(\vec{x})$, and $\sif$ is the standard deviation of $f(\vec{x})$ from its mean.

@@

\par While it is tempting to embrace \eqn{eq:rat1}'s $f({\vec{x}})$ as a nonlinear function and so resort to nonlinear modeling methods, note that a reformulation reveals an underlying linear structure.

@@

  g = \sum_{r=0}^{R} a_r \, \phi_{r}(\vec{x}) + g\, \sum_{v=0}^{V} b_v \, \phi_{v}(\vec{x}) \; .

@@

        b_k, & \quad \;\; \;\text{for }R+1\leq k\leq R+V+1

@@

        \phi_k(\vec{x}_j) g , & \quad \;\; \;\text{for }R+1\leq k\leq R+V+1

@@

\par But note that $\hat{P}$'s depending on $g$ signals that there is more to be done.

@@

In particular, let us briefly consider the effect of noise on $g$, e.g. $g \rightarrow g+n$.

@@

\par The key to robustly solving for $\vec{\alpha}$ lies in iterative refinement.

@@

Specifically, we note that \eqn{eq:rat2} may be modified iteratively minimize the impact of numerical noise on the eventual solution, $\vec{\alpha}$.

@@

To this end, if we define $g^{(0)}$ to be \eqn{eq:rat2} evaluated with $\vec{\alpha}$ as given by \eqn{eq:pinv2}, and note that by construction, $g^{(0)}$ is less effected by (zero mean) noise in $g$, then a further refinement may be obtained by

@@

  g^{(1)} = \sum_{r=0}^{R} a_r \, \phi_{r}(\vec{x}) + g^{(0)}\, \sum_{v=0}^{V} b_v \, \phi_{v}(\vec{x}) \; .

@@

\eqn{eq:rat2} straightforwardly generalizes to

@@

  g^{(n+1)} = \sum_{r=0}^{R} a_r \, \phi_{r}(\vec{x}) + g^{(n)}\, \sum_{v=0}^{V} b_v \, \phi_{v}(\vec{x}) \; .

@@

>>>>

\par The combination of these two ideas alone results in an algorithm prone to a deficit of stepwise methods: the algorithm may confuse correlated basis vectors (e.g. $x^2$ may be confused for $x^4$).

@@

  f(\vec{x}) = \muf + \sif \; \frac{ \sum_{r=0}^{R} a_r \, \phi_{r}(\vec{x}) }{ 1 - \sum_{v=1}^{V} b_v \, \phi_{v}(\vec{x}) } \;,

@@

where $\muf$ is the additive mean of $f(\vec{x})$, and $\sif$ is the standard deviation of $f(\vec{x})$ from its mean, and $\phi_k$ are the multinomials basis functions considered in the previous section.

@@

Note that, in \eqn{eq:rat1}, the sum over $v$ does not include the constant term associated with $\phi_0$.
\par While it is tempting to embrace \eqn{eq:rat1}'s $f({\vec{x}})$ as a nonlinear function and so resort to nonlinear modeling methods, a reformulation reveals an underlying linear structure.

@@

  g = \sum_{r=0}^{R} a_r \, \phi_{r}(\vec{x}) + g\, \sum_{v=1}^{V} b_v \, \phi_{v}(\vec{x}) \; .

@@

        b_k, & \quad \;\; \;\text{for }R+1\leq k\leq R+V

@@

        \phi_k(\vec{x}_j)\, g , & \quad \;\; \;\text{for }R+1\leq k\leq R+V

@@

\par However, we note that $\hat{P}$'s depending on $g$ signals that there is more to be done.

@@

In particular, let us briefly consider the effect of zero-mean noise on $g$, e.g. $g \rightarrow g+n$.

@@

\par The key to robustly solving for $\vec{\alpha}$ lies in the iterative refinement of $\vec{\alpha}$.

@@

Specifically, we note that \eqn{eq:rat2} may be modified to iteratively minimize the impact of numerical noise on $\hat{P}$.

@@

That is, to reduce the impact of noise on $\hat{P}$, we are free to calculate it using model evaluations of $g$ rather than the original training data.
To this end, if we define $g^{(0)}=g$ (i.e. the training data) then, with $\hat{P} = \hat{P}( \vec{g}^{\,(n)} )$, \eqn{eq:pinv2} generalizes to

@@

  \label{eq:pinv3}
  \vec{\alpha}^{(n+1)} = \hat{P}(\vec{g}^{\,(n)}) \, \vec{g}^{(0)}\;.

@@

In practice, one solves \eqn{eq:pinv3} for $\vec{\alpha}^{(n)}$, and then uses the related $a_r^{(n)}$ and $b_v^{(n)}$ to calculate $g^{(n+1)}$ via

@@

  \label{eq:rat4}
  g^{(n+1)} =  \frac{ \sum_{r=0}^{R} a^{(n+1)}_r \, \phi_{r}(\vec{x}) }{ 1\;-\;  \sum_{v=1}^{V} b^{(n+1)}_v \, \phi_{v}(\vec{x}) } \; ,

@@

which is then fed back into \eqn{q:pinv3} for further refinement.
The refinement process is to terminate when a measure of model error (e.g. $||g^{(0)}-g^{(n)}||$) passes a predetermined threshold.
\par Much as in the case of multivariate polynomial fitting, we are left with an un unknown number and content of basis symbols.
In principle, the existence of $a_r$ and $b_v$ makes the problem more complicated, as one could imagine optimizing over each symbol space independently.
To broach these complications, we again use a greedy algorithm with degree tempering.
However, rather than independent greedy optimizations for the numerator and denominator bases symbols, \eqn{eq:rat3} suggests that the appropriate labeling of symbols (e.g. ``numerator'' or ``denominator'') may yield an effective flattening of the supposed 2D basis selection problem.
Put another way, rather than two greedy optimizations over $R$ and $V-1$ symbols (with $(V-1)R$ iterations), a single greedy process over $V+R$ symbols is performed.
\par With the conceptual tools in hand, we may proceed to constructing \gmvr{} be first defining its action, $\mathcal{A}_{\gmvr}$.
This is done in Alg. (\ref{alg:A_gmvr}).
  \caption{$\mathcal{A}_{\gmvr}$, the action for \gmvr. Model calculation given basis symbols, and output of model error estimate.}
  \label{alg:A_gmvr}
